Deep Neural Architectures for Discourse Segmentation in E-Mail Based Behavioral Interventions

### 2018-11-18 paper-amia-summit-2019/manuscript.tex
update manuscript and response to reviewers comments

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include \hl{330 and 281 emails from ecoaches and patients, respectively }(Table~\ref{tab:datastat}). \hl{Each session represents an MI intervention delivered via email. Emails were segmented into 3,138 text fragments and annotated using 115 distinct MYSCOPE}~\cite{carcone2013provider} \hl{behavior codes. We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a class label ``new segment'' or ``same segment'' to indicate whether it precedes a new segment or not. In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.} In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark. Each sample is classified into either a ``new segment'' or a ``same segment'' based on whether there should be a segment break after the current word or not. In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1. \\

\begin{table}[ht]
\centering
\caption{\textbf{\hl{Summary of the experimental dataset}}}
\label{tab:datastat}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class Segments}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
   &  & \textbf{new}  & \textbf{same} & \textbf{word} & \textbf{punctuation}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
 49 & 102,917 & 3,138 & 99,779 & 95,777 & 7,140 & 281 & 330 & MYSCOPE & 115 \\ \hline
  \end{tabular}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\textbf{\hl{Area under the precision-recall curve (AUPR) values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on e-Coaching email segmentation. Highest AUPR value for each feature set across all models is highlighted in boldface.}}}
\label{tab:result_aupr}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\textbf{Features} & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings only & 0.780 & 0.736 & 0.655 & \textbf{0.818} \\ \hline
 word embeddings + POS & 0.797 & 0.746 & 0.647 & \textbf{0.798} \\ \hline
 word embeddings + punctuation & \textbf{0.876} & 0.835 & 0.774 & 0.874 \\ \hline
 all features & \textbf{0.877} & 0.842 & 0.770 & 0.867 \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_aupr} shows the impact of individual features as well as their combination on the segmentation task. \hl{Influence of the punctuation and POS features is also consistent in AUPR values, which increase by 12.44\%, 14.4\%, 17.56\% and 5.99\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together. Individually, POS and punctuation features also improve the performance of all classifiers except BRNN and CRNN when POS features are used. CRF achieved the highest AUPR when punctuation or all features are used together. On the other hand, CRNN demonstrates the highest AUPR when word embeddings are used individually or combined with POS features.}

>>>>

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include \hl{330 and 281 emails from ecoaches and patients, respectively }(Table~\ref{tab:datastat}). \hl{Each session represents an MI intervention delivered via email. Emails were segmented into 3,138 text fragments and annotated using 115 distinct MYSCOPE}~\cite{carcone2013provider} \hl{behavior codes. We consider email segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a class label ``new segment'' or ``same segment'' to indicate whether it precedes a new segment or not. In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances.} In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark. Each sample is classified into either a ``new segment'' or a ``same segment'' based on whether there should be a segment break after the current word or not. In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example is given at $2^{nd}$ row in Table~\ref{tab:datastat}). \\

\begin{table}[ht]
\centering
\caption{\textbf{\hl{Summary of the experimental dataset}}}
\label{tab:datastat}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class Segments}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
   &  & \textbf{new}  & \textbf{same} & \textbf{words} & \textbf{punctuations}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
 49 & 102,917 & 3,138 & 99,779 & 95,777 & 7,140 & 281 & 330 & MYSCOPE & 115 \\ \hline
 \multicolumn{10}{|c|}{segmentation example: Hi[0] XXX[0] ,[0] its[0] good[0] to[0] hear[0] from[0] you[0] .[1] it[0] sounds[0] like[0]...} \\ \hline
  \end{tabular}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\textbf{\hl{Area under the precision-recall curve (AUPR) values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on e-Coaching email segmentation. Highest AUPR value for each feature set across all models is highlighted in boldface.}}}
\label{tab:result_aupr}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\textbf{Features} & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings only & 0.780 & 0.736 & 0.655 & \textbf{0.818} \\ \hline
 word embeddings + POS & 0.797 (+2.18\%) & 0.746 (+1.36\%) & 0.647 (-1.22\%) & \textbf{0.798 (-2.44\%)} \\ \hline
 word embeddings + punctuation & \textbf{0.876 (+12.31\%)} & 0.835 (+13.45\%) & 0.774 (+18.17\%) & 0.874 (+6.85\%) \\ \hline
 all features & \textbf{0.877 (+12.44\%)} & 0.842 (+14.4\%) & 0.770 (+17.56\%) & 0.867 (+6\%) \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_aupr} shows the impact of individual features as well as their combination on the segmentation task. \hl{Influence of the punctuation and POS features is also consistent in AUPR values, which increase by 12.44\%, 14.4\%, 17.56\% and 6\% for CRF, MLP, BRNN and CRNN methods, respectively, when all features are used together. Individually, POS and punctuation features also improve the performance of all classifiers except BRNN and CRNN when POS features are used. CRF achieved the highest AUPR when punctuation or all features are used together. On the other hand, CRNN demonstrates the highest AUPR when word embeddings are used individually or combined with POS features.}
